---
title: "Calibration and Quantification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibration and Quantification}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  message = FALSE,
  warning = FALSE
)
```

```{r}
library(libscanR)
```

## From peak area to concentration

A LIBS calibration converts measured peak area (or intensity ratio) to
elemental concentration using a set of matrix-matched standards.

### Synthetic calibration dataset

```{r data}
cal_ds <- ls_example_data("calibration")
cal_ds
conc <- cal_ds$sample_info$concentration
unique(conc)
```

### Univariate calibration

```{r univariate}
cal <- ls_calibrate(cal_ds, element = "Ca", line_nm = 393.37,
                    concentrations = conc, method = "univariate",
                    verbose = FALSE)
cal
```

### Visualize

```{r plot}
ls_plot_calibration(cal)
```

### Residual diagnostics

```{r residuals}
ls_plot_residuals(cal)
```

### Internal standard calibration

When matrix effects are significant, normalize to a reference line:

```{r internal}
cal_is <- ls_calibrate(cal_ds, "Ca", 393.37, concentrations = conc,
                       method = "internal_std",
                       internal_std_nm = 589.00,
                       verbose = FALSE)
cal_is
```

### PLS calibration

```{r pls, eval = requireNamespace("pls", quietly = TRUE)}
cal_pls <- ls_calibrate(cal_ds, "Ca", 393.37, concentrations = conc,
                        method = "pls", n_components = 5,
                        verbose = FALSE)
cal_pls
```

## Quantifying unknowns

```{r quantify}
ls_quantify(cal, cal_ds)
```

## LOD and LOQ

```{r lod}
ls_lod(cal)
ls_loq(cal)
```

## Calibration-free LIBS (CF-LIBS)

For semi-quantitative estimates without standards, use the Saha-Boltzmann
method:

```{r cf}
spec <- ls_simulate_spectrum(
  elements = c(Ca = 10000, Fe = 500, Na = 2000),
  n_channels = 2048, seed = 7
)
spec <- ls_baseline(spec, method = "snip", iterations = 60)
ls_saha_boltzmann(
  spec, elements = c("Ca", "Fe"),
  lines_nm = list(
    Ca = c(422.673, 445.478, 487.813),
    Fe = c(371.994, 404.581, 438.354)
  ),
  verbose = FALSE
)
```
